---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set( fig.path = "figs/fig-")

```

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", na.strings = "NA")
```

You can see first few rows here
```{r}
head(activity)

```

## What is mean total number of steps taken per day?
Here we first group the data by date. After that we create a histogram of the total number of steps taken each day

```{r}
library(dplyr)
grouped <- group_by(activity, date)
sum_steps <- aggregate(grouped$steps, by = list(activity$date) , FUN = sum, na.rm = T)
names(sum_steps) <- c("date", "total.steps")
hist(sum_steps$total.steps, xlab = "Number of steps", main = "")

```

Here's the **mean** total number of steps taken per day
```{r}
mean(sum_steps$total.steps)
```

Here's the **median** total number of steps taken per day
```{r}
median(sum_steps$total.steps)
```

## What is the average daily activity pattern?
```{r}
mean_steps <- aggregate(grouped$steps, by = list(activity$interval) , FUN = mean, na.rm = T)
plot(mean_steps$Group.1, mean_steps$x, type = "l", xlab = "5-min interval", 
     ylab = "mean number of steps")

```



## Imputing missing values

### Total number of missing values
```{r}
sum(is.na(activity[,1]))
```


### Here we create a function to replace na values with mean value of the interval
```{r}
replace_na <- function(){
    result <- vector()
    for (i in 1:length(activity$steps)){
        if (is.na(activity[i, 1])){
            result <- c(result, mean_steps[mean_steps$Group.1 == activity$interval[i],2])
        } else {
            result <- c(result, activity$steps[i])
        }
    }
    result
}
```

Then we create a new data set, where na values are replaces with the mean value. The new data set is called activity_full

```{r}
activity_full <- activity
activity_full$steps <- replace_na()
```

Here we make a **new** histogram
```{r}
grouped <- group_by(activity_full, date)
sum_steps_full <- aggregate(grouped$steps, by = list(activity_full$date) , FUN = sum)
names(sum_steps_full) <- c("date", "total.steps")
hist(sum_steps_full$total.steps, xlab = "Number of steps", main = "")
```

Before replacing NA values:

mean
```{r}
mean(sum_steps$total.steps)
```
median
```{r}
median(sum_steps$total.steps)
```

After replacing NA values

mean
```{r}
mean(sum_steps_full$total.steps)
```
median
```{r}
median(sum_steps_full$total.steps)
```

There's difference in both mean and median values


## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekend <- c("Saturday", "Sunday")
weekday_type <- function(x){
    if (weekdays(x) %in% weekend){
        return ("Weekend")
    }else {
        return ("Weekday")
    }
}
activity_full$day_type <- as.factor(sapply(as.Date(as.character(activity_full$date)),weekday_type))
```

Making a plot on weekdays/weekends
```{r}
mean_steps <- aggregate(steps ~ interval + day_type, activity_full, mean)
library(ggplot2)
ggplot(mean_steps, aes(x = interval, y = steps)) + 
    geom_line() + facet_wrap(~day_type, ncol = 1) + theme_bw()

```





